---
title: "Day 2 Computational Session"
author: "Jillian Allison"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

rm(list=ls())
```

## Vectors

```{r}
# Create a vector
dogs <- c("teddy","khora","waffle","banjo","tallie") 

# Check the class
class(dogs)
```

```{r}
cats <- c("creamsicle wolf", 14.2, 3L) # going to be character, character is the most flexible type! 

class(cats)
```

```{r}
years <- c(20, 40, 16, 2)

class(years) # returns numeric because we don't have 'L' after each number 

typeof(years) # returns a double - digs deeper into data type 
```

```{r}
# Want to do a conditional check 
# is. <<< asks questions "is this THIS"
# as. <<< gives a command "make this THIS" 

is.numeric(years) # asks if years is numeric --- returns TRUE or FALSE... true in this case. 
is.character(years) # asks if years is a character --- returns FALSE 
as.numeric(years) # coerces years to numeric (even though it already is)
```

## Creating named elements

```{r}
dog_food <- c(teddy = "purina", khora = "alpo", waffle = "fancy feast") 
                    # gives name to each element, stores only values as vector though

dog_food[2] # returns khora "alpo" 

dog_food["khora"] # also returns khora "alpo". calls name associated with value. reproducible!

```

```{r}
cars <- c("blue", "orange", "white")
cars[1]

i <- seq(1:2) # set index value as an object and use that to index
cars[i]

cars[3] <- "yellow" # overwrite existing values
```

## Matrices 

```{r}
fish_size <- matrix(c(12, 14, 18, 20), ncol = 2, nrow = 2, byrow = FALSE)
class(fish_size) # matrix
typeof(fish_size) # double 

# Example matrix indexing 
fish_size[1,1]
fish_size[2,1] # row, column

# What happens if we create a matrix with different types of data? 
fish_type <- matrix(c(12, "trout", 18, 13L), ncol = 2, nrow = 2, byrow = FALSE)
class(fish_type) # matrix
typeof(fish_type) # character

# What happens if I don't specify a row or column? 
fish_size[2,] # gives second row 
fish_size[,2] # gives whole second column 
```

## Lists 

Reminder: A list combines vectors that can be of different length and type

```{r}
urchins <- list("blue", c(1, 2, 3), c("a cat", "a dog"), 5L) 
class(urchins)
typeof(urchins[[1]]) # gives type of each element using DOUBLE BRACKETS! 
```

## Make a named list 

```{r}

tacos <- list(topping = c("onion", "cilantro", "guacamole"),
              filling = c("beans", "meat", "veggies"), 
              price = c(6.75, 8.25, 9.5))
tacos

# Indexing list elements by name 
tacos$filling[2] # searching filling list in tacos -- option 2
```

We can index in this text here!! and have values print!

The second value of car colors is `r cars[2]` USED A BACKTICK TO GET IT IN RMARKDOWN

## Data frames 

A list of vectors where all the vectors have the same length

```{r}
# Creating vectors first, then calling them in data.frame() 
topping <- c("onion", "cilantro", "guacamole")
filling <- c("beans", "meat", "veggies")
price <- c(6.75, 8.25, 9.5)

df_ex_1 <- data.frame(topping, filling, price)

# defining vectors within data.frame()
df_ex_2 <- data.frame(topping = c("sour cream","salsa","cheese"), 
                      filling = c("chicken","cheese","cabbage"),
                      price = c(12.50, 13.75, 15.00))

# Tribble method
df_ex_3 <- tribble(
  ~topping, ~filling, ~price, 
  "avocado", "steak", 13.25, 
  "radish", "fish", 10.40,
  "salsa", "veggies", 9.75
  ) 
```

## Indexing our data frames 

```{r}
df_ex_3$price[3]


```
